A thesis draft

### 2018-11-16 res/sections/25-Tests.tex
Spellcheck and 80 columns

After the testbed implementation, tests about its performance and metrics were
gathered in order to understand the real project effectiveness. We performed
tests relatively two main sections: one testing the architecture overall
responsivness and one checking the SFC implementation and efficiency.

\section{System responsivness}

@@

Regarding the overall system responsivness, a metric we considered worth to 
test was the VNF start up time in a docker environment versus one virtualized 
through a common virtual machine system (in this case Virtual Box was choosen). 
To make this simulation as fair as possible, we used the same node (an 
Openstack VM vith 32GB of RAM, 8 vCPU and solid state storage) and used the same 
``boot sequence'' for both the VNFs: first we launched the Astaire framework, 
ensuring its availability to process data and then we notified the test 
machine of the successful boot. The startup time was measured from the 
beginning of the startup command (\verb!docker run! for Docker and 
\verb!VBoxManage -s! for Virtual Box) till the first TCP hit received by the 
test backend (to make things as smooth as possible, \verb!netcat! was used 
to listen to incoming TCP data).

It is worth saying that Docker metrics are generally more precise than the
Virtual Box one: Docker allows to inspect the container and to gain startup and
shutdown times with a precision of millseconds, while to gather this information
for VirtualBox instances we had to use the command line utility \verb!time!,
that registered timing with a precision of hundredths of a second. Nonetheless,
this didn't influence the test too much. To reduce possible outliers we perfomed
the test $100$ times. With this $N$, the total time spent by the Virtual
Machines to boot up was of $2021,15$ seconds, while with containers the same
number of start ups resulted in a total of $1.638$ seconds. Summing up, Docker
employs only the $0,08\%$ of the overall time spent for the test, underlining
how isolating the process with a set of defined policy and avoiding a
virtualized kernel boot-up saves a considerable amount of time.

@@

\begin{table}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|c|c}
                     & \textbf{Standard deviation (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
\textbf{Docker}      & 0.009620874                           & 0,016                                    \\
\textbf{Virtual Box} & 0.755                                 & 20,211                                 
\end{tabular}%
}
\caption[Docker vs Virtual Box start up times comparison]{Docker vs Virtual Box
  start up times comparison. Is possible to se how the overall time required to
  start $100$ containers is less than the time to start one traditional virtual
  machine. The overall time required by the traitional virtualization system is
  approximately of 33 minutes, witch gives a significant insight of how much
  time is required to create and boot a virtualized enviroment.}
\label{my-label}
\end{table}

@@

\paragraph*{Kubernetes specifications}
A Kubernetes installation was performed in order to carry out the test
descripted above. All the machines involved (master node and minions) had the
same specifications i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage.
Nonetheless the constrained resources, absolutely far from a real Kubernetes
production environment, it is possible to notice how, even with a chain of ten
elements, the average time required to deploy an SFC is equal, if not smaller,
of the average time required to deploy a Virtual Box machine on an instance 
with 32GB of RAM, 8 vCPU and 50GB of SSD storage.

@@

\subsection{RTT measurement}
Regarding the delivery performance, a metric to take in account is the \emph{round trip time}(RTT), that is the time interval between the instant in whic a packet is sent and the response is received. To test it there was developed a script that sends a packet with 8B of payload and waits for the corrispondant response, recording the instants (expressed as \emph{epoch}) of send and receive. The sake of thi test is to compare the RTT measured without any SFC (only sending the packet to the receiver waiting for the response) and making packets pass through different length SFCs, in particular with 1, 2, 3, 4, 5, and 10 hops. Packets traverse the SFC on both sides and the results are collected after 100 tests and the SF only prints the packet received and the total number of packets that have passed throught it.

@@

Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
the RTT increase with the number of elements of the chain. There is a big difference between the mean RTT of the configuration wiothout any chain and the one with an SFC with a single link: this outcome has different explaination. As explained in \todo{Add reference to implementation} the chain has always at least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3 different running software. Another explaination lays into how the code was developed: it was not developed as a production ready product, but for a feasibility test, so connections and trasmissions can be enhanced. Finally, SFs are meant to perform some modifications to the packet, so, unavoidably, some time is waisted.

@@

The results are better represented in Figure~\ref{chap:tests:sec:rtt:img:meanstt}. In that graph we can easier see the difference among the chains with different lengths and the trasmission that does not use any SFC. Looking at the bars representing the SFC composed by 1, 2, 3, 4 and 5 links we can see a pretty linear increase of the RTT. This trend strongly depends on the nature of the SF involved in the SFC, in that case the linear increase is highligted due to the use of the same funcion. In Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference between the RTT measured with a normal packet transmission and using the chains with different size, that represent the additional time needed for each transmission. Even in that case we can see that some time is waisted against a normal transmission and that this delta of time depends on the number of SFC involved in the chain.

@@

In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the difference in terms of RTT among chains that have a length difference of one link. The first bar represent the comparison between an SFC composed by only one element and the direct trasmission to the receiver: that explain why that value is higher than the others. Other differences among SFC, instead, are quite similar.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[]
\centering
\begin{tabular}{@{}cc@{}}
\toprule
\textbf{SFC length} & \textbf{Jitter (Milliseconds)} \\ \midrule
len. 0              & 0.0001074767    \\
len. 1              & 0.0028695822    \\
len. 2              & 0.0032037902    \\
len. 3              & 0.0031961584    \\
len. 4              & 0.0027278513    \\
len. 5              & 0.0036802184    \\
len. 10             & 0.0063434506    \\ \bottomrule
\end{tabular}
\caption{Jitterf for different SFC lengths}
\label{chap:tests:sec:rtt:tab:jitter}
\end{table}

Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for different SFC length, based on values of RTT measured in the previous test. Jitter is the variation in the delay of received packet, and gives a measure of the congestion of the network. As we can see values are really low, even if the lowest one is always the one calculated without the chain.

>>>>

After the testbed implementation, tests about its performance and metrics were
gathered in order to understand the real project effectiveness. We performed
tests relatively two main sections: one testing the architecture overall
responsiveness and one checking the SFC implementation and efficiency.

\section{System responsiveness}

@@

Regarding the overall system responsiveness, a metric we considered worth to 
test was the VNF start up time in a docker environment versus one virtualized 
through a common virtual machine system (in this case Virtual Box was chosen). 
To make this simulation as fair as possible, we used the same node (an 
Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the same 
``boot sequence'' for both the VNFs: first we launched the Astaire framework, 
ensuring its availability to process data and then we notified the test 
machine of the successful boot. The startup time was measured from the 
beginning of the startup command (\verb!docker run! for Docker and 
\verb!VBoxManage -s! for Virtual Box) till the first TCP hit received by the 
test backend (to make things as smooth as possible, \verb!netcat! was used 
to listen to incoming TCP data).

It is worth saying that Docker metrics are generally more precise than the
Virtual Box one: Docker allows to inspect the container and to gain startup and
shutdown times with a precision of millseconds, while to gather this information
for VirtualBox instances we had to use the command line utility \verb!time!,
that registered timing with a precision of hundredths of a second. Nonetheless,
this didn't influence the test too much. To reduce possible outliers we performed
the test $100$ times. With this $N$, the total time spent by the Virtual
Machines to boot up was of $2021,15$ seconds, while with containers the same
number of start ups resulted in a total of $1.638$ seconds. Summing up, Docker
employs only the $0,08\%$ of the overall time spent for the test, underlining
how isolating the process with a set of defined policy and avoiding a
virtualized kernel boot-up saves a considerable amount of time.

@@

\begin{table}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{l|c|c}
                     & \textbf{Standard deviation (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
\textbf{Docker}      & 0.009620874                           & 0,016                                    \\
\textbf{Virtual Box} & 0.755                                 & 20,211                                 
\end{tabular}%
}
\caption[Docker vs Virtual Box start up times comparison]{Docker vs Virtual Box
  start up times comparison. Is possible to see how the overall time required to
  start $100$ containers is less than the time to start one traditional virtual
  machine. The overall time required by the traditional virtualization system is
  approximately of 33 minutes, witch gives a significant insight of how much
  time is required to create and boot a virtualized environment.}
\label{my-label}
\end{table}

@@

\paragraph*{Kubernetes specifications}
A Kubernetes installation was performed in order to carry out the test
described above. All the machines involved (master node and minions) had the
same specifications i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage.
Nonetheless the constrained resources, absolutely far from a real Kubernetes
production environment, it is possible to notice how, even with a chain of ten
elements, the average time required to deploy an SFC is equal, if not smaller,
of the average time required to deploy a Virtual Box machine on an instance 
with 32GB of RAM, 8 vCPU and 50GB of SSD storage.

@@

\subsection{RTT measurement}
Regarding the delivery performance, a metric to take in account is the
\emph{round trip time}(RTT), that is the time interval between the instant in
which a packet is sent and the response is received. To test it there was
developed a script that sends a packet with 8B of payload and waits for the
correspondent response, recording the instants (expressed as \emph{epoch}) of
send and receive. The sake of this test is to compare the RTT measured without
any SFC (only sending the packet to the receiver waiting for the response) and
making packets pass through different length SFCs, in particular with 1, 2, 3,
4, 5, and 10 hops. Packets traverse the SFC on both sides and the results are
collected after 100 tests and the SF only prints the packet received and the
total number of packets that have passed through it.

@@

Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
the RTT increase with the number of elements of the chain. There is a big
difference between the mean RTT of the configuration without any chain and the
one with an SFC with a single link: this outcome has different explanation. As
explained in \todo{Add reference to implementation} the chain has always at
least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3
different running software. Another explanation lays into how the code was
developed: it was not developed as a production ready product, but for a
feasibility test, so connections and transmissions can be enhanced. Finally, SFs
are meant to perform some modifications to the packet, so, unavoidably, some
time is wasted.

@@

The results are better represented in
Figure~\ref{chap:tests:sec:rtt:img:meanstt}. In that graph we can easier see
the difference among the chains with different lengths and the transmission that
does not use any SFC. Looking at the bars representing the SFC composed by 1,
2, 3, 4 and 5 links we can see a pretty linear increase of the RTT. This trend
strongly depends on the nature of the SF involved in the SFC, in that case the
linear increase is highlighted due to the use of the same function. In
Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference
between the RTT measured with a normal packet transmission and using the chains
with different size, that represent the additional time needed for each
transmission. Even in that case we can see that some time is wasted against a
normal transmission and that this delta of time depends on the number of SFC
involved in the chain.

@@

In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the
difference in terms of RTT among chains that have a length difference of one
link. The first bar represent the comparison between an SFC composed by only
one element and the direct transmission to the receiver: that explain why that
value is higher than the others. Other differences among SFC, instead, are
quite similar.

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[]
\centering
\begin{tabular}{@{}cc@{}}
\toprule
\textbf{SFC length} & \textbf{Jitter (Milliseconds)} \\ \midrule
len. 0              & 0.0001074767    \\
len. 1              & 0.0028695822    \\
len. 2              & 0.0032037902    \\
len. 3              & 0.0031961584    \\
len. 4              & 0.0027278513    \\
len. 5              & 0.0036802184    \\
len. 10             & 0.0063434506    \\ \bottomrule
\end{tabular}
\caption{Jitter for different SFC lengths}
\label{chap:tests:sec:rtt:tab:jitter}
\end{table}

Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for
different SFC length, based on values of RTT measured in the previous test.
Jitter is the variation in the delay of received packet, and gives a measure of
the congestion of the network. As we can see values are really low, even if the
lowest result is always the one calculated without the chain.
